---
title: "Query openrouteservice from R"
author: "Andrzej Oleś"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
vignette: >
  %\VignetteIndexEntry{Query openrouteservice from R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r config, include=FALSE}
## set up knitr defaults
knitr::opts_chunk$set(eval=TRUE, out.width='100%', out.height='560px')
```

## Get started

<!-- README START -->
```{r doc, include=FALSE}
## create alias
doc <- openrouteservice:::doc_link
```

The *openrouteservice* package provides easy access to the
[openrouteservice](https://openrouteservice.org) (ORS) API from R. 
It allows you to painlessly consume the following services:

  - `r doc('directions')`
  - `r doc('geocode')` powered by [Pelias](https://pelias.io)
  - `r doc('isochrones')`
  - `r doc('matrix')`
  - `r doc('pois')`

### Disclaimer

By using this package, you agree to the ORS [terms and
conditions](https://openrouteservice.org/terms-of-service/).

### Installation

The package is not yet available from CRAN, but you can install the
development version directly from GitHub.

```{r installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("GIScience/openrouteservice-r")
```
<!-- README END -->

### Setting up API key

In order to start using ORS services you first need to set up your personal API key, which you can 
[get for free](https://openrouteservice.org/sign-up/).

```{r api_key, eval=FALSE}
library(openrouteservice)

ors_api_key("<your-api-key>")
```

This will save the key in the default keyring of your system credential store. Once the key is defined, it persists in the keyring store of the operating system. This means that it survives beyond the termination of the R session, so you don't need to set it again each time you start a new R session. To retrieve the key just call `ors_api_key()` without the `key` argument.


## Directions

`ors_directions()` interfaces the ORS directions service to compute routes between given `coordinates`. 

```{r directions, eval=TRUE}
library(openrouteservice)

coordinates = list(c(8.34234, 48.23424), c(8.34423, 48.26424))

x = ors_directions(coordinates, format="geojson")
```

By setting `format="geojson"` the result can formatted as geoJSON which allows to easily [visualize](https://rstudio.github.io/leaflet/json.html#working-with-raw-geojsontopojson) it with [leaflet](https://CRAN.R-project.org/package=leaflet). 

```{r leaflet}
library(leaflet)

leaflet() %>%
  addTiles() %>%
  addGeoJSON(x, fill=FALSE) %>%
  fitBBox(x$bbox)
```

The API offers a wide range of `profile`s for multiple modes of transport: car, different bicycle types, walking, hiking, wheelchair and heavy vehicle. Each of these modes uses a carefully compiled street network to suite the profiles requirements.

```{r bicycle}
x = ors_directions(coordinates, profile="cycling-mountain", elevation=TRUE, format="geojson")

leaflet() %>%
  addTiles() %>%
  addGeoJSON(x, fill=FALSE) %>%
  fitBBox(x$bbox)
```

Setting `elevation = TRUE` adds height to the coordinates of the points along the route, which can be used to plot the elevation profile.

```{r elevation}
library(sp)

## create data.frame
coords = x$features[[1]]$geometry$coordinates
colnames(coords) = c("Longitude", "Latitude", "Height")

dists = spDists(coords[,c("Longitude", "Latitude")], longlat=TRUE, segments=TRUE)

plot(c(0, cumsum(dists)), coords[, "Height"], type = "l", main = "Elevation profile", xlab = "Distance", ylab = "Height")
```


## Isochrones

Reachability has become a crucial component for many businesses from all different kinds of domains. `ors_isochrones()` helps you to determine which areas can be reached from certain location(s) in a given time or travel distance. The reachability areas are returned as contours of polygons. Next to the `range` provided in seconds or meters you may as well specify the corresponding `interval`s. If the input entails a list of points the service will return an intersections of the computed isochrones if any exist. `ors_isochrones()` accepts identical arguments as `ors_directions()`.

```{r isochrones}
coordinates = list(c(8.34234, 48.23424), c(8.34234, 47.23424))

x = ors_isochrones(coordinates, range = 60*60, interval = 20*60)

# set isochrones color
ranges = x$info$query$ranges
pal = setNames(heat.colors(length(ranges)), ranges)

x$features <- lapply(1:length(x$features), function(i) {
  feature = x$features[[i]]
  range = feature$properties$value

  ## set style
  col = unname(pal[as.character(range)])
  feature$properties$style <- list(color = col, fillColor = col, fillOpacity=0.5)

  ## restrict polygon to current level only
  if (range > ranges[1])
    feature$geometry$coordinates = c(feature$geometry$coordinates, x$features[[i-1]]$geometry$coordinates)

  feature
})

leaflet() %>%
  addTiles() %>%
  addGeoJSON(x) %>%
  fitBBox(x$bbox)
```


## Matrix

One to many, many to many or many to one. `ors_matrix()` allows you to obtain aggregated time and distance information between a set of locations (origins and destinations). Unlike `ors_directions()` it does not return detailed route information. But you may still specify the transportation mode and compute routes which adhere to certain restrictions, such as avoiding specific road types or object characteristics.

```{r matrix}
coordinates = list(
  c(9.970093, 48.477473),
  c(9.207916, 49.153868),
  c(37.573242, 55.801281),
  c(115.663757,38.106467)
)

# query for duration and distance in km
res = ors_matrix(coordinates, metrics = c("duration", "distance"), units = "km")

# duration in hours
(res$durations / 3600) %>% round(1)

# distance in km
res$distances
```

## Geocoding

`ors_geocode()` transforms a description of a location provided in `query`, such as the place's name, street address or postal code, into a normalized description of the location with a point geometry. Additionally, it offers reverse geocoding which does exactly the opposite: It returns the next enclosing object which surrounds the coordinates of the given `location`. To obtain more relevant results you may also set a radius of tolerance around the requested coordinates. 

```{r geocode}
## locations of Heidelberg around the globe
x = ors_geocode("Heidelberg")

leaflet() %>%
  addTiles() %>%
  addGeoJSON(x) %>%
  fitBBox(x$bbox)

## set the number of results returned
x = ors_geocode("Heidelberg", size = 1)

## search within a particular country
x = ors_geocode("Heidelberg", boundary.country = "DE")

## structured geocoding
x = ors_geocode(list(locality="Heidelberg", county="Heidelberg"))

## reverse geocoding
location = x$features[[1L]]$geometry$coordinates

y = ors_geocode(location = location, layers = "locality", size = 1)
```


## POIs

This service allows you to find places of interest around or within given geographic coordinates. You may search for given features around a point, path or even within a given polygon specified in `geometry`. To list all the available POI categories use `ors_pois('list')`.

```{r pois}
geometry = list(
  geojson = list(
    type = "Point",
    coordinates = c(8.8034, 53.0756)
  ),
  buffer = 500
)

ors_pois(request = 'pois',
         geometry = geometry,
         limit = 2000,
         sortby = "distance",
         filters = list(
           category_ids = I(488),
           wheelchair = I("yes")
         ))
```

You can gather statistics on the amount of certain POIs in an area by using `request='stats'`.

```{r stats}
ors_pois(request = 'stats',
         geometry = geometry,
         limit = 2000,
         sortby = "distance",
         filters = list(
           category_ids = I(488)
         ))
```
